Supplementary Information

### 2020-06-07 doc/manuscript/MEE_manuscript.Rmd
worked on manuscript

**Characterizing the influence of slowly-changing environtmental drivers is very challenging and uncertain.**
We want to know how changes in CO~2~ and atmospheric deposition are influencing tree growth.
A signficant challenge arises when tree-rings are used to characterize the influence of slowly changing environmental variables--climate, CO2, atmospheric deposition--on tree growth rates (REFS).
In these cases, a variety of methods are used to account for simultaneous changes in tree size, including... (DENDRO_REFS).
One commonly applied approach is to assume that $BAI$ is fundamentally independent of size (DENDRO_REFS).
None of these methods is fully satisfactory, leaving great uncertainty as to what the tree-ring record can tell us about the impact of rising CO~2~ (Walker et al. in revision). 
Rather, it is necessary to *simultaneously* account for the influences of changing tree size and environmental drivers.

**Here, develop a new method that allows simultaneous consideration of the effects of tree size, objectively determined principle climate drivers, and other environemntal drivers on annual tree growth.** 
This allows us to ask:
(1) What are the most important climate drivers of annual growth, and over which time windows?
(2) What is the shape of the relationship between annual growth and climate drivers?
(3) How do $\Delta r$, $BAI$, and aboveground biomass increments ($\Delta AGB$) vary with DBH?
(4) Are there interactions between DBH and climate drivers common?
(5) How is growth simultaneously influenced by DBH, climate, and other environemental drivers (CO~2~, atmospheric deposition)?

@@

We then used the *climwin* package in R (REF) [(van de Pol et. al. 2016)](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12590) to identify the most significant climate driver and time window for each of the three climate variable groups. 
*Climwin* searches through multiple climate variables and over a wide range of time frames to identify which are most strongly correlated to the biological variable of interest--in this case, residuals of $log[\Delta r]$, $log[BAI]$, or $log[\Delta AGB]$.
Within *climwin*, we specified a mixed effects model using species and tree identity as random effect: `residual ~ [climate] + (1 | sp) + (1 | treeID)`. 
Here, for each permutation, `climate` specifies one of the climate drivers in the climate variable group, analyzed over one of all possible combinations of time periods, at monthly resolution, starting in the previous May and ending in August of the current year.
We note that analyzing all species together yields the most significant climate drivers across the full set of cores from each site (our goal here), whereas identification of the top climate drivers for individual species would be optimized by analyzing each species separately.
*Climwin* can check for linear and quadratic relationships, and does k-fold cross-validation in its computation of AIC to guard against over-fitting (CLIMWIN_REF). 
We specified quadratic relationships only because (i) quadratic relationships are more consistent with known biological mechanisms (see Introduction), (ii) preliminary tests revealed that quadratic fits usually had lower AIC, and when not there tended to be little difference in AIC and the curve would approximate a straight line, (iii) modeling only quadratic improves coding efficiency and speeds up the analysis process.
For each group of candidate climate variables, we move forward with the best variable over the time window identified by climwin as a candidate climate variable for the multivariate models. 
When PET came out as the best variable in both temperature and cloud cover groups, there were only two candidate variables for the GAM. 
We checked for colinearity among the selected climate variables and removed any variable with a variance inflation factor > 10.

@@

We get different climate senstivity results with the different metrics of growth.

**Figure/ Table on climate - DBH interactions?**

@@

**Figure showing effects of CO2, SO2, NOx**

>>>>

**Characterizing the influence of slowly-changing environtmental drivers is very challenging and uncertain.** (*perhaps save for Bianca's paper? See [ISSUE #41 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/41)*)
We want to know how changes in CO~2~ and atmospheric deposition are influencing tree growth.
A signficant challenge arises when tree-rings are used to characterize the influence of slowly changing environmental variables--climate, CO2, atmospheric deposition--on tree growth rates (REFS).
In these cases, a variety of methods are used to account for simultaneous changes in tree size, including... (DENDRO_REFS).
One commonly applied approach is to assume that $BAI$ is fundamentally independent of size (DENDRO_REFS).
None of these methods is fully satisfactory, leaving great uncertainty as to what the tree-ring record can tell us about the impact of rising CO~2~ (Walker et al. in revision). 
Rather, it is necessary to *simultaneously* account for the influences of changing tree size and environmental drivers.

**Here, develop a new method that allows simultaneous consideration of the effects of tree size, objectively determined principle climate drivers, and other environemntal drivers on annual tree growth.** 
This allows us to ask:
(1) What are the most important climate drivers of annual growth, and over which time windows?
(2) What is the shape of the relationship between annual growth and climate drivers?
(3) How do $\Delta r$, $BAI$, and aboveground biomass increments ($\Delta AGB$) vary with DBH?
(4) Are there interactions between DBH and climate drivers common?
(5) How is growth simultaneously influenced by DBH, climate, and other environemental drivers (CO~2~, atmospheric deposition)? (*perhaps save 5 for Bianca's paper? See [ISSUE #41 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/41)*)

@@

We then used the *climwin* package in R (REF) [(van de Pol et. al. 2016)](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12590) to identify the most significant climate driver and time window for each of the three climate variable groups. 
*Climwin* searches through multiple climate variables and over a wide range of time frames to identify which are most strongly correlated to the biological variable of interest--in this case, residuals of $log[\Delta r]$, $log[BAI]$, or $log[\Delta AGB]$.
Within *climwin*, we specified a mixed effects model using species and tree identity as random effect: `residual ~ [climate] + (1 | sp) + (1 | treeID)`. 
Here, for each permutation, `climate` specifies one of the climate drivers in the climate variable group, analyzed over one of all possible combinations of time periods, at monthly resolution, starting in the previous May and ending in August of the current year.
We note that analyzing all species together yields the most significant climate drivers across the full set of cores from each site (our goal here), whereas identification of the top climate drivers for individual species would be optimized by analyzing each species separately.
*Climwin* can check for linear and quadratic relationships, and does k-fold cross-validation in its computation of AIC to guard against over-fitting (CLIMWIN_REF). 
We specified quadratic relationships only because (i) quadratic relationships are more consistent with known biological mechanisms (see Introduction), (ii) preliminary tests revealed that quadratic fits usually had lower AIC, and when not there tended to be little difference in AIC and the curve would approximate a straight line, (iii) modeling only quadratic improves coding efficiency and speeds up the analysis process.
For each group of candidate climate variables, we move forward with the best variable over the time window identified by climwin as a candidate climate variable for the multivariate models. 
When PET came out as the best variable in both temperature and cloud cover groups, there were only two candidate variables for the GAM. 
If PET came out as the best variable in one but not both of these groups, we compared the AICs of the top variables from each group, retaining only the stronger predictor.
We checked for colinearity among the selected climate variables and removed any variable with a variance inflation factor > 10.

@@

We get different climate senstivity results with the different metrics of growth (Fig. 5?).
Preliminary review of some early results indicates that the "landscape" of climate effects over various time windows is usually(?) similar across the three metrics of growth, but that the optimal time window or even top climate variable can shift.
(**See [ISSUE #40 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/40)**).
The underlying cause of these differences would be differences in the variance structure, where growth residuals would tend to be greater (in absolute value) when growth rate is high. Thus, use of $\Delta r$ as a growth metric would tend to place proportionaly more weight on smaller $DBHs$, whereas $\Delta AGB$ would tend to place more weight on larger individuals, with $BAI$ internediate (Fig. 4). (**Christy, please check/revise**)
The degree to which climate senstivity landscapes differ across growth metrics should depend on the size structure of the data and the existence/strength of climate-DBH interactions.

![**Figure 5 | Comparison of climwin output by growth metric:** Current example is PET at Lilly Dickey, selected mainly because data/model for other sites aren't yet stable. (I'm not sure if this is interesting enough to present in the main text. If not, we should have at least one example in the SI)](tables_figures/climwin_by_growth_metric.png)

**Table on climate - DBH interactions? (KAT needs to discuss with VH)**

@@

**Figure showing effects of CO2, SO2, NOx** (*perhaps save for Bianca's paper? See [ISSUE #41 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/41)*)
